---
title: "Fuqing Exercise  4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
setwd('/Users/gfq/Desktop/Genomic data sciences/HW4/Exercise 4-20221120')
```

### Exploring the count tables

Here, import an example count table and do some exploration of the expression data. 

```{r exSetup1, eval=FALSE}
counts_file <- system.file("extdata/rna-seq/SRP029880.raw_counts.tsv", package = "compGenomRData")
coldata_file <- system.file("extdata/rna-seq/SRP029880.colData.tsv",
                            package = "compGenomRData")
```

1. Normalize the counts using the TPM approach. 
```{r, eval=FALSE}
counts <- as.matrix(read.table(counts_file, header = T, sep = '\t'))

geneLengths <- as.vector(subset(counts, select = c(width)))

# Computing TPM
#find gene length normalized values
rpk <- apply( subset(counts, select = c(-width)), 2,
              function(x) x/(geneLengths/1000))

#normalize by the sample size using rpk values
tpm <- apply(rpk, 2, function(x) x / sum(as.numeric(x)) * 10^6)

# Check the sample sizes of tpm. Notice that the sums of samples are all equal to 10^6.
colSums(tpm)

```

2. Plot a heatmap of the top 500 most variable genes. Compare with the heatmap obtained using the 100 most variable genes. Hint : use var() function to compute the variance. 
```{r, eval=FALSE}
library(compGenomRData)
library(pheatmap)
library(stats)
library(corrplot)

V <- apply(tpm, 1, var)
selectedGenes <- names(V[order(V, decreasing = T)][1:500])

#sort the results by variance in decreasing order
#and select the top 500 genes

colData <- read.table(coldata_file, header = T, sep = '\t',
                      stringsAsFactors = TRUE)
pheatmap(tpm[selectedGenes,], scale = 'row',
         show_rownames = FALSE,
         annotation_col = colData)
```

3. Re-do the heatmaps setting the `scale` argument to `none`, and `column`. Compare the results with `scale = 'row'`.
```{r, eval=FALSE}
pheatmap(tpm[selectedGenes,], scale = 'none',
         show_rownames = FALSE,
         annotation_col = colData)

pheatmap(tpm[selectedGenes,], scale = 'column',
         show_rownames = FALSE,
         annotation_col = colData)
```

4. Draw a correlation plot for the samples depicting the sample differences as 'ellipses', drawing only the upper end of the matrix, and order samples by hierarchical clustering results based on `average` linkage clustering method. 

```{r, eval=FALSE}

# Correlation plot
correlationMatrix <- cor(tpm)

# Dissimilarity matrix
d <- dist(correlationMatrix, method = "euclidean")

# Hierarchical clustering using Average Linkage
hc1 <- hclust(d, method = 'average' )
plot(hc1)

corrplot(correlationMatrix[1:5,1:5], order = 'hclust',method = 'ellipse',
         addrect = 2, addCoef.col = 'white',
         number.cex = 0.7)

```

### Git and Github

5. Open Github account, and install git if it required as instructed in the following link: https://github.com/llendway/github_for_collaboration/blob/master/github_for_collaboration.md

6. Create a repository and commit and push the answers of this exercise through Rstudio (Please upload the answers to the exercises also to Moodle).

7. Insert the link to your Github account here.
```{r, eval=FALSE}
https://github.com/Olliegfq/Fuqing_test_repo.git
```

## Test